---
title: "Loan Status Prediction"
output: html_notebook
---

## Load Libraries

```{r message=F, echo=F}
library(tidyverse)
library(DT)
library(data.table)
```

## Data Inspection and Cleaning

```{r}
loan_Df <- read.csv("C:/Users/PC/Documents/R_4DS/bank_loan.csv", na.strings = c("", " ", NA))

glimpse(loan_Df)
```

Glimpse of dataset indicates an approach to Supervised Learning, a couple of explanatory variablesinfluencing whether or not a Loan is availed, a binarized outcome implying a Logistic Regression approach to understanding these underlying/obvious variable relationship.

```{r}
## Check Missing Values
null_vars <- (sapply(loan_Df, function(x) sum(is.na(x))))
t(data.frame(null_vars))
```

- Having a *Missing* credit history is as good as having None, will change NAs here to 0
- Will not hurt to make the missing Loan Amount the mean of all known requested Loan
- ..


```{r}
## Blank Rows
blank_rows <- (sapply(loan_Df, function(x) sum(x == "")))
t(data.frame(blank_rows))
```


```{r}
loan_Df <- loan_Df %>% 
  mutate(Credit_History = replace_na(Credit_History, mean(Credit_History, na.rm = T))) %>% 
  mutate(LoanAmount = replace_na(LoanAmount, mean(LoanAmount, na.rm = T))) %>% 
  mutate(Dependents = replace_na(Dependents, mean(Dependents, na.rm = T))) %>% 
  mutate(Loan_Amount_Term = replace_na(Loan_Amount_Term, min(Dependents, na.rm = T))) %>% 
  fill(Gender, .direction =  'downup') %>% 
  fill(Self_Employed, .direction =  'updown') %>% 
  filter(!Married == "") %>% 
  mutate(Credit_History = ifelse(Credit_History < 1, 0, Credit_History)) %>% 
  mutate(Loan_Amount_Term = case_when(Loan_Amount_Term <= 120 ~ "Short_Termed",
                                    Loan_Amount_Term <= 300 ~ "Mid_Termed",
                                    Loan_Amount_Term > 300 ~ "Long_Termed")) %>%  
  mutate(Loan_Amount_Term = factor(Loan_Amount_Term, levels = c("Short_Termed", "Mid_Termed", "Long_Termed"))) %>% 
  select(-c(Loan_ID))
```


## Exploratory Data Analysis

```{r}
## Distrubtution of the Response Variable
loan_Df %>% 
  ggplot(aes(x = Loan_Status))+
  geom_bar()
```

Over response data is imbalanced.

## Modelling

#### Preprocessing Data
```{r}
library(mltools)
ml_df <- data.table(loan_Df[-12])

ml_df_encoded <- one_hot(ml_df)
ml_df_encoded <- cbind(ml_df_encoded, loan_Df[12])
quick_sample(ml_df_encoded)
```

```{r}
### Train-Test
n_train <- round(0.8*nrow(ml_df_encoded))
train_indices <- sample(1:nrow(ml_df_encoded), n_train)
df_train <- ml_df_encoded[train_indices,]
df_test <- ml_df_encoded[-train_indices,]
```

```{r}
### Logistic
logi_model <- glm(formula = Loan_Status ~ Gender_Female + Gender_Male + 
                   Married_Yes + Dependents + Education_Graduate + 
                   Self_Employed_No + ApplicantIncome + CoapplicantIncome + 
                   LoanAmount + Loan_Amount_Term_Short_Termed + 
                   Loan_Amount_Term_Mid_Termed + Loan_Amount_Term_Long_Termed + 
                   Credit_History + Property_Area_Rural + 
                   Property_Area_Semiurban + Property_Area_Urban, family = "binomial", data = df_train)

#disable scientific notation for model summary
options(scipen=999)

#view model summary
summary(logi_model)
```


## Evaluation
```{r}
library(pscl)

pR2(logi_model)["McFadden"]
```

A value of **0.1838804** is low for McFadden’s R2, which indicates that our model does not fit the data very well and has low predictive
power.

```{r}
## VIP Variables
library(caret)

vip_vars <- data.frame(varImp(logi_model))
vip_vars %>% 
  arrange(-Overall)
```

```{r}
library(InformationValue)
## Make Predictions
predicted <- predict(logi_model, df_test, type="response")

optimal <- optimalCutoff(df_test$Loan_Status, predicted)[1]
optimal
```
#+> prediction from a rank-deficient fit may be misleading
This tells us that the optimal probability cutoff to use is **0.1185697**. Thus, any individual with a probability of defaulting of **0.1185697** or higher will be predicted to get the Loan, while any individual with a probability less than this number will be predicted to not be availed a Loan.

```{r}
## Checking
confusionMatrix(df_test$Loan_Status, predicted)
```

```{r}
#calculate sensitivity
sensitivity(df_test$Loan_Status, predicted)

#calculate specificity
specificity(df_test$Loan_Status, predicted)

#calculate total misclassification error rate
misClassError(df_test$Loan_Status, predicted, threshold=optimal)
```

