---
title: "Logistic Regression of Bank Loans"
author: "AfroLogicInsect_MLReview"
date: "4/29/2022"
output: html_document
---

## Load Library
```{r output = FALSE, echo = FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(caret)
library(data.table)
library(ROSE)
library(caTools)
library(pROC)
library(PRROC)
library(mvtnorm)
```


## Load Dataset
```{r}
loan_Df <- read.csv("C:/Users/PC/Documents/R_4DS/bank_loan.csv", stringsAsFactors = FALSE)
```


## Exploratory Dataset
```{r}
##drop cases where there is no Loan Amount
df <- loan_Df %>% 
  filter_at(vars(LoanAmount, Loan_Amount_Term), all_vars(!is.na(.))) %>% 
  mutate(Loan_Status = if_else(Loan_Status == "Y", 1,0)) %>% 
  mutate(Gender = if_else(Gender == "Male", 1,0)) %>%
  mutate(Married = if_else(Married == "Yes",1,0)) %>%
  mutate(Self_Employed = if_else(Self_Employed == "Yes",1,0)) %>%
  mutate(Education = if_else(Education == "Graduate", 1,0)) %>%
  mutate(Property_Area = if_else(Property_Area == "Rural", 0, if_else(Property_Area == "Semiurban", 1, 2))) %>% 
  select(-c(Loan_ID))

df$Credit_History[is.na(df$Credit_History)] <- 0
```
Dropped `r dim(loan_Df)[1] - dim(df)[1]` rows of data that provided no sense to being empty, if the **Credit History** is NULL then it is better as `0` and dropped the LoanID column for now.

```{r}
## Visualizing
### Checking for Bias in the Loan Status: Dependent Variable

library(scales)
plotdata <- count(loan_Df, Loan_Status) %>% 
  mutate(pct = n/sum(n),
         pctlabel = paste0(round(pct*100), "%"))


plotdata

plotdata %>% 
  ggplot(aes(x = reorder(Loan_Status, -pct),
             y = pct)) +
  geom_bar(stat = "identity", fill = "indianred3", color = "black") + 
  geom_text(aes(label = pctlabel),
            vjust=-0.25) +
  scale_y_continuous(labels = percent) +
  labs(x = "Status", y = "Frequency", title = "Loan Status: Check for Bias")

```


```{r}
# ggplot(loan2 , aes(x = grade , y = int_rate , fill = grade)) + 
#         geom_boxplot() + 
#         theme_igray() + 
#         labs(y = 'Interest Rate' , x = 'Grade')

# ggplot(loan2[sample(244179 , 10000) , ] , aes(x = annual_inc , y = loan_amnt , color = int_rate)) +
#         geom_point(alpha = 0.5 , size = 1.5) + 
#         geom_smooth(se = F , color = 'darkred' , method = 'loess') +
#         xlim(c(0 , 300000)) + 
#         labs(x = 'Annual Income' , y = 'Loan Ammount' , color = 'Interest Rate')
```




Most of the transactions were Approved for Loan (`r plotdata$pctlabel[plotdata$Loan_Status == "Y"]`) of the time, while Declined Loans occurs (`plotdata$pctlabel[plotdata$Loan_Status == "N"]`) of the time in the dataframe. This presents us with an IMBALANCE.

## Modeling
```{r}
##Convert Dependent Variable to Numeric
##Train-Test Split

n_train <- round(0.8 * nrow(df))
train_indices <- sample(1:nrow(df), n_train)
df_train <- df[train_indices, ]
df_test <- df[-train_indices, ] 
```


```{r}
## Feature Scaling
training_set <- df_train 
test_set <- df_test  

###---|| NB: We do not scales the Response Variable;Data has to be numeric.
training_set[-12] = scale(training_set[-12])
test_set[-12] = scale(test_set[-12])
```

###Apply Logistic classifier on Imbalanced data

```{r}
# Apply Logistic classifier on training set
normal_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = training_set)
# Predicting the test set using Under sampling classifier
normal_probability_predict = predict(normal_classifier, type = 'response', newdata = test_set[-12])
normal_probability_predict = zoo::na.aggregate(normal_probability_predict)
y_pred_normal = ifelse(normal_probability_predict>0.5, 1, 0)

# To check the model accuracy using confusionMatrix
confusionMatrix(table(test_set[, 12], y_pred_normal))

# To check the accuracy of this model using ROC curve.
roc_over <- PRROC::roc.curve(test_set$Loan_Status, y_pred_normal, curve = TRUE)
plot(roc_over)
```
Error in if (auc < 0.5) { : missing value where TRUE/FALSE needed

#### Logistic Regression Modelling with Balanced Dataset

```{r SmoteFunction}
## SMOTE Function
SMOTE <- function(form,data,
                  perc.over=200,k=5,
                  perc.under=200,
                  learner=NULL,...
)
  
  # INPUTS:
  # form a model formula
  # data the original training set (with the unbalanced distribution)
  # minCl  the minority class label
  # per.over/100 is the number of new cases (smoted cases) generated
  #              for each rare case. If perc.over < 100 a single case
  #              is generated uniquely for a randomly selected perc.over
  #              of the rare cases
  # k is the number of neighbours to consider as the pool from where
  #   the new examples are generated
# perc.under/100 is the number of "normal" cases that are randomly
#                selected for each smoted case
# learner the learning system to use.
# ...  any learning parameters to pass to learner
{
  
  # the column where the target variable is
  tgt <- which(names(data) == as.character(form[[2]]))
  minCl <- levels(data[,tgt])[which.min(table(data[,tgt]))]
  
  # get the cases of the minority class
  minExs <- which(data[,tgt] == minCl)
  
  # generate synthetic cases from these minExs
  if (tgt < ncol(data)) {
    cols <- 1:ncol(data)
    cols[c(tgt,ncol(data))] <- cols[c(ncol(data),tgt)]
    data <-  data[,cols]
  }
  newExs <- smote.exs(data[minExs,],ncol(data),perc.over,k)
  if (tgt < ncol(data)) {
    newExs <- newExs[,cols]
    data <- data[,cols]
  }
  
  # get the undersample of the "majority class" examples
  selMaj <- sample((1:NROW(data))[-minExs],
                   as.integer((perc.under/100)*nrow(newExs)),
                   replace=T)
  
  # the final data set (the undersample+the rare cases+the smoted exs)
  newdataset <- rbind(data[selMaj,],data[minExs,],newExs)
  
  # learn a model if required
  if (is.null(learner)) return(newdataset)
  else do.call(learner,list(form,newdataset,...))
}



# ===================================================
# Obtain a set of smoted examples for a set of rare cases.
# L. Torgo, Feb 2010
# ---------------------------------------------------
smote.exs <- function(data,tgt,N,k)
  # INPUTS:
  # data are the rare cases (the minority "class" cases)
  # tgt is the name of the target variable
  # N is the percentage of over-sampling to carry out;
  # and k is the number of nearest neighours to use for the generation
  # OUTPUTS:
  # The result of the function is a (N/100)*T set of generated
  # examples with rare values on the target
{
  nomatr <- c()
  T <- matrix(nrow=dim(data)[1],ncol=dim(data)[2]-1)
  for(col in seq.int(dim(T)[2]))
    if (class(data[,col]) %in% c('factor','character')) {
      T[,col] <- as.integer(data[,col])
      nomatr <- c(nomatr,col)
    } else T[,col] <- data[,col]
  
  if (N < 100) { # only a percentage of the T cases will be SMOTEd
    nT <- NROW(T)
    idx <- sample(1:nT,as.integer((N/100)*nT))
    T <- T[idx,]
    N <- 100
  }
  
  p <- dim(T)[2]
  nT <- dim(T)[1]
  
  ranges <- apply(T,2,max)-apply(T,2,min)
  
  nexs <-  as.integer(N/100) # this is the number of artificial exs generated
  # for each member of T
  new <- matrix(nrow=nexs*nT,ncol=p)    # the new cases
  
  for(i in 1:nT) {
    
    # the k NNs of case T[i,]
    xd <- scale(T,T[i,],ranges)
    for(a in nomatr) xd[,a] <- xd[,a]==0
    dd <- drop(xd^2 %*% rep(1, ncol(xd)))
    kNNs <- order(dd)[2:(k+1)]
    
    for(n in 1:nexs) {
      # select randomly one of the k NNs
      neig <- sample(1:k,1)
      
      ex <- vector(length=ncol(T))
      
      # the attribute values of the generated case
      difs <- T[kNNs[neig],]-T[i,]
      new[(i-1)*nexs+n,] <- T[i,]+runif(1)*difs
      for(a in nomatr)
        new[(i-1)*nexs+n,a] <- c(T[kNNs[neig],a],T[i,a])[1+round(runif(1),0)]
      
    }
  }
  newCases <- data.frame(new)
  for(a in nomatr)
    newCases[,a] <- factor(newCases[,a],levels=1:nlevels(data[,a]),labels=levels(data[,a]))
  
  newCases[,tgt] <- factor(rep(data[1,tgt],nrow(newCases)),levels=levels(data[,tgt]))
  colnames(newCases) <- colnames(data)
  newCases
}
```


```{r}
library(DMwR2)
library(ROSE)

## -> For some reason the SMOTE function defined above would not work without the Response Variable as Factor
training_set$Loan_Status <- as.factor(training_set$Loan_Status)

print('Number of transactions in train dataset before applying sampling methods')
print(table(training_set$Loan_Status))

# Oversampling, as Fraud transactions(1) are having less occurrence, 

over_sample_train_data <- ovun.sample(Loan_Status ~ ., data = training_set, method="over", N=456)$data
print('Number of transactions in train dataset after applying Over sampling method')
print(table(over_sample_train_data$Loan_Status))

# Undersampling,
#But, you see that we’ve lost significant information from the sample. 
under_sample_train_data <- ovun.sample(Loan_Status ~ ., data = training_set, method="under", N=228)$data
print('Number of transactions in train dataset after applying Under sampling method')
print(table(under_sample_train_data$Loan_Status))

# Mixed Sampling, apply both under sampling and over sampling on this imbalanced data
both_sample_train_data <- ovun.sample(Loan_Status ~ ., data = training_set, method="both", p=0.5, seed=222, N=228)$data
print('Number of transactions in train dataset after applying Mixed sampling method')
print(table(both_sample_train_data$Loan_Status))

# ROSE Sampling, this helps us to generate data synthetically. It generates artificial datas instead of dulicate data.
rose_sample_train_data <- ROSE(Loan_Status ~ ., data = training_set,  seed=111)$data
print('Number of transactions in train dataset after applying ROSE sampling method')
print(table(rose_sample_train_data$Loan_Status))

# SMOTE(Synthetic Minority Over-sampling Technique) Sampling
# formula - relates how our dependent variable acts based on other independent variable.
# data - input data
# perc.over - controls the size of Minority class
# perc.under - controls the size of Majority class
# since my data has less Majority class, increasing it with 200 and keeping the minority class to 100.
smote_sample_train_data <- SMOTE(Loan_Status ~ ., data = training_set, perc.over = 100, perc.under=200)
print('Number of transactions in train dataset after applying SMOTE sampling method')
print(table(smote_sample_train_data$Loan_Status))
```

```{r}
##Apply Logistic classifier on balanced data

# Now we have five different types of inputs which are balanced and ready for prediction.
# We can appply Logistic classifier to all these five datasets and calculate the performance of each.

# Logistic classifier for Over sampling dataset
over_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = over_sample_train_data)

# Logistic classifier for Under sampling dataset
under_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = under_sample_train_data)

# Logistic classifier for Mixed sampling dataset
both_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = both_sample_train_data)

#Logistic classifier for ROSE sampling dataset
rose_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = rose_sample_train_data)

# Logistic classifier for SMOTE dataset
smote_classifier = glm(formula = Loan_Status ~ ., family = binomial, data = smote_sample_train_data)
```

```{r}
# Predicting the test set using Over sampling classifier
over_probability_predict = predict(over_classifier, type = 'response', newdata = test_set[-12])
y_pred_over = ifelse(over_probability_predict>0.5, 1, 0)

# Predicting the test set using Under sampling classifier
under_probability_predict = predict(under_classifier, type = 'response', newdata = test_set[-12])
y_pred_under = ifelse(under_probability_predict>0.5, 1, 0)

# Predicting the test set using Mixed sampling classifier
both_probability_predict = predict(both_classifier, type = 'response', newdata = test_set[-12])
y_pred_both = ifelse(both_probability_predict>0.5, 1, 0)

# Predicting the test set using ROSE classifier
rose_probability_predict = predict(rose_classifier, type = 'response', newdata = test_set[-12])
y_pred_rose = ifelse(rose_probability_predict>0.5, 1, 0)

# Predicting the test set using SMOTE classifier
smote_probability_predict = predict(smote_classifier, type = 'response', newdata = test_set[-12])
y_pred_smote = ifelse(smote_probability_predict>0.5, 1, 0)
```


## Evaluating Model (with ROC Curve)


```{r}
library(patchwork)
library(cowplot)

test_set$Loan_Status <- as.factor(test_set$Loan_Status)
# For reson(s) this keeps generating NA* values
# y_pred_both = zoo::na.aggregate(y_pred_both) 
# y_pred_normal = zoo::na.aggregate(y_pred_normal) 
# y_pred_over = zoo::na.aggregate(y_pred_over) 
# y_pred_rose = zoo::na.aggregate(y_pred_rose) 
# y_pred_smote = zoo::na.aggregate(y_pred_smote) 
# y_pred_under = zoo::na.aggregate(y_pred_under) 

pred_list <- c("y_pred_both", "y_pred_normal", "y_pred_over", "y_pred_rose", "y_pred_smote", "y_pred_under")

for(pred in pred_list) assign(pred, zoo::na.aggregate(get(pred)))

layout(matrix(c(1,1,2,3), 2, 2, byrow = TRUE))
# ROC curve of over sampling data
roc_over <- PRROC::roc.curve(test_set$Loan_Status, y_pred_over, curve = T)
 
plot(roc_over,
main="ROC Curve: Original Distribution")

# ROC curve of Under sampling data
roc_under <- PRROC::roc.curve(test_set$Loan_Status, y_pred_under, curve = T)

plot(roc_under,
main="ROC Curve: UnderSampled Distribution")

# ROC curve of both sampling data
roc_both <- PRROC::roc.curve(test_set$Loan_Status, y_pred_both, curve = T)
plot(roc_both,
main="ROC Curve: Mixed-Sampled Distribution")

# ROC curve of Random Over Sampling Examples [ROSE]sampling data
roc_rose <- PRROC::roc.curve(test_set$Loan_Status, y_pred_rose, curve = T)
roc_rose_p <- plot(roc_rose,
main="ROC Curve: ROSE Distribution")

# ROC curve of Synthetic Minority Over-Sampling Technique [SMOTE] sampling data
roc_smote <- PRROC::roc.curve(test_set$Loan_Status, y_pred_smote, curve = T)
roc_smote_p <- plot(roc_smote,
main="ROC Curve: SMOTEDistribution")
 
# roc_under_p + roc_over_p + roc_both_p + roc_rose_p + roc_smote_p + plot_layout(ncol = 1, heights = c(1, 5))
```

```{r}
library(pROC)

roc_under <- roc(test_set$Loan_Status, y_pred_under)
ggroc(roc_under)

# figurw
```

